---
title: "Review of Wash Trading Methodologies"
author: "charliemarketplace.eth"
date: '2022-10-17'
output:
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

This markdown seeks to compare and contrast open source methodologies
for identification and exclusion of "wash" trade transactions of NFTs.

It makes no judgement of the methodologies nor implies their accuracy in
isolation or against each other. Wash-trading is difficult to identify
and NFT trading is a highly adversarial environment.

It will focus specifically on the count and type of transactions
exclusions across different filters; and the marginal effect of
excluding transactions on various common NFT trading statistics.

There is no target dataset available that perfectly labels wash-trades,
thus metrics such as accuracy, precision, sensitivity, and specificity
will not be discussed here.

# Data

The following NFT trade histories are collected from the Flipside Crypto
EZ_NFT_Sales tables for the Ethereum blockchain. Specification on why
the collection was included are also provided.

These 2 collections will be loosely referenced as "benchmark"
collections because they have trades on all 3 platforms.

-   Bored Ape Yacht Club: Top traded collection with volume on OpenSea,
    LooksRare, and X2Y2, the 3 exchanges of interest for this analysis.
-   Azuki: A Top traded collection on all three of OpenSea, LooksRare,
    and X2Y2.

These 3 collections will be loosely referred to as "test" collections
because they are more concentrated in some exchanges versus others;
and/or have had large historical shifts on which exchange they trade on.

-   Meebits: High volume collection with a large shift in primary
    platform traded; originally high volume disproportionally on
    LooksRare, but shifting to X2Y2 and OpenSea.
-   More Loot: A high volume collection on LooksRare and X2Y2, with
    significantly lower volume on OpenSea.
-   RENGA: A high volume collection on OpenSea and X2Y2, significantly
    lower on LooksRare.

All NFT collections can have wash-trading. The goal of this analysis is
to quantify the impacts of transaction exclusion based on different open
source methodologies including impact on particular exchanges and
collections which have different websites, users, communities, and fees.

Data is capped at a block height of: 15760000 for reproducibility.

```{r, warning = FALSE, message = FALSE}
library(shroomDK)
library(reactable)
library(plotly)
library(dplyr)

query <- {
"SELECT BLOCK_NUMBER, BLOCK_TIMESTAMP, TX_HASH, PLATFORM_NAME, SELLER_ADDRESS, 
BUYER_ADDRESS, NFT_ADDRESS, PROJECT_NAME, TOKENID, CURRENCY_SYMBOL, 
PRICE, PRICE_USD, PLATFORM_FEE, CREATOR_FEE 
FROM ethereum.core.ez_nft_sales 
WHERE NFT_ADDRESS IN (
'0xbc4ca0eda7647a8ab7c2061c2e118a18a936f13d', --bayc 
  '0xed5af388653567af2f388e6224dc7c4b3241c544', -- azuki
  '0x7bd29408f11d2bfc23c34f18275bbf23bb716bc7', -- meebits
  '0x1dfe7ca09e99d10835bf73044a23b73fc20623df', -- More Loot
  '0x394e3d3044fc89fcdd966d3cb35ac0b32b0cda91' -- RENGA
) AND BLOCK_NUMBER <= 15760000
ORDER BY BLOCK_NUMBER ASC
"
}

sales <- shroomDK::auto_paginate_query(query, api_key = readLines("api_key.txt"))
sales$PROJECT_NAME[is.na(sales$PROJECT_NAME)] <- 'renga' # not labeled in data 
```

# Methodology

Two methods from [hildobby's DUNE
Spellbook](https://github.com/duneanalytics/spellbook/pull/1623) are
considered:

-   **D**une **S**ame **T**raders in last **7** days (dst7): Exclude
    transactions where the buyer & seller exchanged the same NFT within
    7 days of the transaction.
-   **D**une **S**ame **B**uyer **3** times (dsb3): Exclude transactions
    where the buyer has previously purchased the same NFT 3 or more
    times ever.

Three methods from a private DUNE dashboard, not linked here (yet).

-   **S**ame **D**ay **s**ame **P**latform **3** times (sdsp3): Exclude
    transactions where the same NFT was traded 3+ times on the same
    platform on the same day.

-   **S**ame **D**ay, **S**ame **T**raders **S**ame **P**latform **3**
    times (sdstsp4): Exclude transactions where the buyer & seller
    exchanged any NFTs in the same collection 3+ times on the same platform on the same
    day.

-   **P**rice **10x** in last **7** Days (p10x7): Exclude transactions
    where the sale price is \>50x the maximum sale price of the NFT in
    the last 7 days.

To review these methods the following columns are added to the data:

-   day\_: day of sale as a date
-   traders: an *alphabetized* concatenation of BUYER_ADDRESS and
    SELLER_ADDRESS; such that the value is the same for the same buyer
    and seller pair no matter who direction the trade went (i.e., Bob
    sells to Alice and Alice sells to Bob are both always Alice-Bob) to
    create a unique id for a pair of traders.
-   token_traders: The concatenation of NFT Address, Token ID, and traders.

```{r}

sales <- sales %>%
  mutate(day_ = as.Date(BLOCK_TIMESTAMP))

# speed optimization
traders <- character(nrow(sales)) 
token_traders <- character(nrow(sales))

for(i in 1:nrow(sales)){
  traders[i] <- paste0( 
    sort( 
      c(sales$BUYER_ADDRESS[i], sales$SELLER_ADDRESS[i]) 
    ),
    collapse = "-")
  
  token_traders[i] <-
    paste0(
      c(sales$NFT_ADDRESS[i], 
        sales$TOKENID[i],
        traders[i]),
    collapse = "-")
  
}

sales$traders <- traders
sales$token_traders <- token_traders

rm(traders)
rm(token_traders)

```

Implementing each method as a TRUE/FALSE column for whether the transaction meets the condition.

dst7: Dune Same Traders in Last 7 Days.

Group buy NFT, TOKENID, and Trader Pair; Arrange by day of sale, and if the last time
these traders traded the same NFT was within 7 days, mark that transaction as dst7 = TRUE.

```{r}

sales <- sales %>% group_by(token_traders) %>%
  arrange(token_traders, day_) %>% 
  mutate(dst7 = ifelse( day_ - lag(day_) <= 7, TRUE, FALSE) )

sales$dst7[is.na(sales$dst7)] <- FALSE

```

dsb3: Dune Same Buyer 3+ Times

Group by NFT, TOKENID, and BUYER; Arrange by day of sale; if this same buyer has bought this 
same NFT 3+ times *ever*, mark the 3rd through Nth purchases as dsb3 = TRUE. 

```{r}
sales <- sales %>% 
  group_by(NFT_ADDRESS, TOKENID, BUYER_ADDRESS, .add = TRUE) %>% 
  arrange(NFT_ADDRESS, TOKENID, BUYER_ADDRESS, day_) %>% 
  mutate(dsb3 = ifelse(row_number() >= 3, TRUE, FALSE)) 

```

sdsp3: Same Day Same Platform 3+ Times 

Group by NFT, TOKENID, Platform and Day of Sale; if the same NFT traded on the same day on 
the same platform 3+ times; mark *all* of those sales as sdsp3 = TRUE.

```{r}
sales <- sales %>% 
    group_by(NFT_ADDRESS, TOKENID, PLATFORM_NAME, day_, .add = TRUE) %>% 
  mutate(sdsp3 = ifelse(n() >= 3, TRUE, FALSE)) 

```

sdstsp3: Same Day, Same Trader, Same Platform 3+ Times

GROUP BY NFT, Platform, Traders, and Day: If the same Traders traded the same collection on the same platform 3+ times on the same day, mark all their trades of that NFT collection as sdstsp3 = TRUE.

```{r}
sales <- sales %>% 
    group_by(NFT_ADDRESS, traders, PLATFORM_NAME, day_, .add = TRUE) %>% 
  mutate(sdstsp3 = ifelse(n() >= 3, TRUE, FALSE)) 

```

p10x7: Price 10x+ of highest sale price in collection of last 7 days.

GROUP BY NFT, Order by Day, track max PRICE_USD (although most trades are in ETH or WETH, some are not, filter to ETH/WETH trades to use PRICE in ETH terms) in last 7 days, and if the price is 10x or larger than max price in last 7 days, mark p10x7 = TRUE.

```{r}
sales <- sales %>% 
  mutate(day_num = as.numeric(day_))

p10x7 <- unique(
  as.data.frame(sales[, c("NFT_ADDRESS", "day_num")])
)

get_max <- function(address_day_row, trades_data){

  nft = address_day_row[[1]]
  the_day = as.numeric(address_day_row[[2]])
  
  r = NA 
  
  temp_tbl <- trades_data %>% filter(
    NFT_ADDRESS == nft, 
    day_num > the_day - 7, 
    day_num < the_day
  )
  if(nrow(temp_tbl) != 0){
  r = max(temp_tbl["PRICE_USD"])
  }
  
    return(r)
}
  
p10x7$max_price <- apply(X = p10x7, MARGIN = 1,
                         FUN = get_max, 
                         trades_data = as.data.frame(sales))

sales <- merge(sales, p10x7, all.x = TRUE, all.y = TRUE,
                        by.x = c("NFT_ADDRESS","day_num"), by.y = c("NFT_ADDRESS","day_num"))

sales <- sales %>% 
  mutate(p10x7 = ifelse(PRICE_USD >= 10*max_price, TRUE, FALSE))

sales$max_price <- NULL
sales$p10x7[is.na(sales$p10x7)] <- FALSE 

sales$num_flags <- apply(sales[, c("dst7", "dsb3", "sdsp3","sdstsp3","p10x7")], 
                        MARGIN = 1, 
                        FUN = sum)

```

# Results by Collection

The benchmark collections: Azuki and Bored Ape Yacht Club exist across

```{r}

as_percent <- function(x){
  paste0(round(x),"%")
}

sales_tbl <- sales %>% group_by(PROJECT_NAME) %>% 
  summarise(num_sales = n(),
            num_dst7 = sum(dst7),
            percent_dst7 = as_percent(sum(dst7)*100/n()),
            num_dsb3 = sum(dsb3),
            percent_dsb3 = as_percent(sum(dsb3)*100/n()),
            num_sdsp3 = sum(sdsp3),
            percent_sdsp3 = as_percent(sum(sdsp3)*100/n()),
            num_sdstsp3 = sum(sdstsp3),
            percent_sdstsp3 = as_percent(sum(sdstsp3)*100/n()),
            num_p10x7 = sum(p10x7),
            percent_p10x7 = as_percent(sum(p10x7)*100/n()),
            num_1plus_flag = sum(num_flags > 0),
            percent_1plus_flag = as_percent(sum(num_flags > 0)*100/n())
  ) %>% as.data.frame()

colnames(sales_tbl) <- c("Project", "# Sales",
                         "# DST7", "DST7 %",
                         "# DSB3", "DSB3 %",
                         "# SDSP3", "SDSP3 %",
                         "# SDSTSP3", "SDSTSP3 %",
                         "# p10x7", "p10x7 %",
                         "# 1+ Metric", "1+ Metric %")

reactable(sales_tbl)

```

# Results by Platform 

```{r}

platform_tbl <- sales %>% group_by(PLATFORM_NAME) %>% 
  summarise(num_sales = n(),
            num_dst7 = sum(dst7),
            percent_dst7 = as_percent(sum(dst7)*100/n()),
            num_dsb3 = sum(dsb3),
            percent_dsb3 = as_percent(sum(dsb3)*100/n()),
            num_sdsp3 = sum(sdsp3),
            percent_sdsp3 = as_percent(sum(sdsp3)*100/n()),
            num_sdstsp3 = sum(sdstsp3),
            percent_sdstsp3 = as_percent(sum(sdstsp3)*100/n()),
            num_p10x7 = sum(p10x7),
            percent_p10x7 = as_percent(sum(p10x7)*100/n()),
            num_1plus_flag = sum(num_flags > 0),
            percent_1plus_flag = as_percent(sum(num_flags > 0)*100/n())
  ) %>% as.data.frame()

colnames(platform_tbl) <- c("Platform", "# Sales",
                         "# DST7", "DST7 %",
                         "# DSB3", "DSB3 %",
                         "# SDSP3", "SDSP3 %",
                         "# SDSTSP3", "SDSTSP3 %",
                         "# p10x7", "p10x7 %",
                         "# 1+ Metric", "1+ Metric %")

reactable(platform_tbl)
```

# Comparison of Metric Results 

## DST7 - the most aggressive method

Surprisingly, on the margins the most aggressive metric was DST7: The same 
buyer and seller trading the same NFT TOKEN ID within the last 7 days.

This is *not* to say this metric was accurate in judging a transaction as a true 
positive sale. 

It is simply noting that this metric consistently labeled the largest amount 
of transactions while also capturing a significant portion of all other metrics 
labels as well.

## SDSP3 and SDSTP3 were exactly 100% correlated

Note: Among the 5 NFT collections, there is a 100% correlation between metrics 
SDSP3 and SDSTSP3. That is; in *every* instance that the same NFT ID was traded
on the same day on the same platform 3+ times (SDSP3); it was *always* done 
by the **same traders** at the collection level. The corollary is also true, that 
the same traders trading the same NFT *collections* on the same day on the same platform, *always* traded the same TOKENIDs 3+ times as well. 

There was not a single instance of traders balancing the counts of trading 
token IDs while trading within a collection, nor traders triangulating TOKENID 
trades between a 3rd/4th/5th address.

## P10X7 is an extremely passive label.

Trades occuring at 10x+ the highest USD price in the last 7 days (not including 
the same day) were extremely rare (< 1% of trades). In situations where this happens, it is possible that specific traits of the NFT (i.e., rarity) are the cause. 

# Conclusion 

With p10x7 being so rarely applied; and sdsp3 & sdstsp3 being 100% correlated. 
Comparing dst7, dsb3, and sdsp3 as the metrics of most marginal information 
shows the following correlation matrix.

**Without information on false positives and false negatives** these metrics
cannot be optimized toward a target rate.

```{r}

plot_ly(z = cor(sales[,19:21]), 
        x = ~colnames(sales[,19:21]), 
        y = ~colnames(sales[,19:21]), 
        type = 'heatmap') %>% 
  layout(xaxis = list(title = ""),
         yaxis = list(title = "")
         )

```
